---
title: "BiocBBSpack: build system concepts in R for Bioconductor"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{BiocBBSpack overview}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

Building and distributing packages for Linux, macOS, and Windows in a continuous
integration framework is a hallmark of the Bioconductor project.  The
Bioconductor Build System has evolved over years and has proven
robust and resilient as the project has grown in scope, and as
the underlying languages and infrastructures have evolved.

The package BiocBBSpack is an exploration of how recently developed
tools in R and Bioconductor can contribute to the performance and
maintainability of the build system.

# Simple example

With an example we illustrate some of the most basic concepts.

## The PackageSet class

```{r lklk}
suppressPackageStartupMessages({
library(BiocBBSpack)
})
coreset() # a simple collection of relatively lighteight packages
ps1 = PackageSet(coreset())
ps1
ps1 = add_dependencies(ps1)
ps1
```

The large number of dependencies for
this small group of packages is characteristic.
One reason for working on this package is to provide
a framework for experimenting with package management
processes to measure scalability.  We would like to
pass the responsibility for efficient installation
of essential packages to R or Bioconductor utilities
that are designed for this purpose.  In this case,
our current understanding is that the installation of
dependencies can be parallelized by passing a suitable
value for parameter `Ncpus` to `install.packages` via
`BiocManager::install`.

## Populating a folder with git clones of sources in a PackageSet

```{r dopop}
tf = tempfile()
dir.create(tf)
populate_local_gits(ps1, tf)
dir(tf)
```

For a large collection of packages this can be time-consuming but
can be regarded as a one-time-charge, because updates can be
conducted as needed via `git pull`.

## Identifying an out-of-date source clone

Here we artificially back-version a package and
show how to find it, comparing its version value
to that in the Bioconductor repository.  This task could be carried out
at the git level, but that can occur in a more mature
version if necessary.
```{r lkout}
okdesc = readLines(paste0(tf, "/parody/DESCRIPTION"))
baddesc = gsub("Version.*", "Version: 1.0", okdesc)
writeLines(baddesc, paste0(tf, "/parody/DESCRIPTION"))
local_gits_behind_bioc(tf)
writeLines(okdesc, paste0(tf, "/parody/DESCRIPTION")) # restore
```

## Summary of simple example

- We use a character vector of package names to
specify the build task.  The `PackageSet` class
manages relevant information about these packages:
    - package names,
    - dependencies as identified via Depends, Imports, 
LinkingTo, or Suggests; the latter are essential for 
building vignettes,
    - additional information as returned by `BiocPkgTools::biocPkgList()`
- We use the concepts of `gits` and `gitspath` to refer to
collections of cloned package source repositories held in a single
folder.
    - `populate_local_gits` will attempt to clone sources into
a specified folder,
    - `local_gits_behind_bioc` will list packages for which the
version specified in DESCRIPTION is earlier than the one reported
in `BiocPkgTools::biocPkgList()`.

# Conceptual framework

Now that we have seen a little bit of the functionality
of BiocBBSpack, let's try to spell out the concepts
that must be covered in a build system support package.

## Generating current source set

- `base source provider`: a collection of git repositories owned
by Bioconductor
- `package set`: a list of packages and metadata, to be managed with `PackageSet`
class
- `gitspath`: a folder where all repositories corresponding to 
elements of a package set are cloned; these clones are `local_gits`
- `lagging_local_gits`: clones whose version tag in DESCRIPTION
is earlier than the corresponding one in the base source provider
- `make_all_current_in_gitspath`: task of running `git pull` so that
there are no `lagging_local_gits`

## Provisioning builder infrastructure

- `ubuntu package list`: list of arguments to apt-get install to
achieve a capable build machine
- `manual infrastructure spec`: directions to add infrastructure (typically
runtime libraries and header sets) to achieve a capable build machine
- `full CRAN package set`: `PackageSet` of all CRAN packages needed
for a build
- `current-full-R`: installation of R with all packages required to
build any Bioconductor package installed

## Generating logs and tarballs 

Package build attempts may fail, and logs must be kept and conveyed.
When build succeeds, logs and tarballs are generated

- `current log-tarball set`: convergent result of iteration
of the build system -- we iterate because a given attempt may
fail owing to resource contention or other random challenges
- `lagging-log-tarballs`: logs and tarballs for builds of source
packages whose version is earlier than that at the base source provider

## Generating binaries

- `current binary set`: binary builds of packages from `current log-tarball set`
- `lagging-binaries`: binary builds with versions earlier than that at
the base source provider

<!--
> ls("package:BiocBBSpack")
 [2] "bioc_software_packagelist"               
 [3] "build1"                                  
 [4] "build1_with_buildsink"                   
 [6] "default_packtxt"                         
 [7] "endow_ubuntu"                            
 [8] "full_dep_opts"                           
 [9] "getpk"                                   
[10] "installed_r"                             
[11] "list_packs_to_update"                    
[12] "local_gits_behind_bioc"                  
[13] "local_gits_with_uninstalled_dependencies"
[14] "PackageSet"                              
[15] "parallel_tarballs"                       
[16] "populate_local_gits"                     
-->
